Highly Available Transactions: Virtues and Limitations

### 2013-03-19 writeup/model.tex
more writing

\begin{figure}
\centering
\begin{tikzpicture}[scale=0.8]
  \tikzstyle{every node}=[font=\small]
 \node[draw=none,fill=none] (1) at (0,0) {$1$-availability}; 
 \node[draw=none,fill=none] (2) at (0,1) {$2$-availability}; 
 \node[draw=none,fill=none] (m-1) at (0, 2) {\ldots}; 
 \node[draw=none,fill=none] (m) at (0, 3) {majority availability}; 
 \node[draw=none,fill=none] (m+1) at (0, 4) {\ldots}; 
 \node[draw=none,fill=none] (n) at (0, 5) {$N$-availability}; 
 \node[draw=none,fill=none] (1s) at (4, 1) {$1$-sticky availability};
 \node[draw=none,fill=none] (2s) at (4, 2) {$2$-sticky availability};
 \node[draw=none,fill=none] (n2s) at (4, 3) {\ldots};
 \node[draw=none,fill=none] (n1s) at (4, 4) {$(N$$-$$1)$-sticky availability};

 \draw [->] (1) -- (2);
 \draw [->] (2) -- (m-1);
 \draw [->] (m-1) -- (m);
 \draw [->] (m) -- (m+1);
 \draw [->] (m+1) -- (n);
 \draw [->] (1) -- (1s);
 \draw [->] (1s) -- (2s);
 \draw [->] (2) -- (2s);
 \draw [->] (2s) -- (n2s);
 \draw [->] (n2s) -- (n1s);
 \draw [->] (n1s) -- (n);

@@

We show a hierarchy of replica availabilty levels in
Figure~\ref{fig:availability-order}. $K$-availability is ``more
available'' than $K$-sticky available, but, in general, $K$-sticky
availability is incomparable with $(K+1)$-sticky
availability. However, $N$-sticky availability is effectively the same
as $N$-availability--if operations must be performed on all replicas,
then subsequent requests will trivially access the same set of
replicas.

>>>>

\begin{figure}
\centering
\begin{tikzpicture}[scale=0.8]
  \tikzstyle{every node}=[font=\small]
 \node[draw=none,fill=none] (1) at (0,0) {$1$-availability (high availability)}; 
 \node[draw=none,fill=none] (2) at (0,1) {$2$-availability}; 
 \node[draw=none,fill=none] (m-1) at (0, 2) {\ldots}; 
 \node[draw=none,fill=none] (m) at (0, 3) {majority availability}; 
 \node[draw=none,fill=none] (m+1) at (0, 4) {\ldots}; 
 \node[draw=none,fill=none] (n) at (0, 5) {$N$-availability}; 
 \node[draw=none,fill=none] (1s) at (4, 1.5) {$1$-sticky availability};
 \node[draw=none,fill=none] (2s) at (4, 2.5) {$2$-sticky availability};
 \node[draw=none,fill=none] (n2s) at (4, 3.5) {\ldots};
 \node[draw=none,fill=none] (n1s) at (4, 4.5) {$(N$$-$$1)$-sticky availability};

 \draw [->] (1) -- (2);
 \draw [->] (2) -- (m-1);
 \draw [->] (m-1) -- (m);
 \draw [->] (m) -- (m+1);
 \draw [->] (m+1) -- (n);
 \draw [->] (1) -- (1s);
 \draw [->] (1s) -- (m);
 \draw [->] (2s) -- (m+1);
 \draw [->] (1s) -- (2s);
 \draw [->] (2) -- (2s);
 \draw [->] (2s) -- (n2s);
 \draw [->] (n2s) -- (n1s);
 \draw [->] (n1s) -- (n);

@@

We show a hierarchy of replica availabilty levels in
Figure~\ref{fig:availability-order}. $K$-availability is ``more
available'' than $K$-sticky available, but, in general, $K$-sticky
availability is incomparable with $(K+1)$-sticky
availability. However, $N$-sticky availability is effectively the same
as $N$-availability--if operations must be performed on all replicas,
then subsequent requests will trivially access the same set of
replicas. Similarly, with majority availability, at least one replica
will be present in all requests, so $1$-sticky availability is weaker
than majority availability, and, more generally, $K$-sticky
availability is weaker than $\lceil \frac{N}{2} \rceil$$+$$K$$-$$1$
availability.

### 2013-03-19 writeup/motivation.tex

Peter Deutsch begins his list of ``Fallacies of Distributed
Computing'' with two concerns: ``\textit{1.)}  The network is
reliable. \textit{2.)} Latency is zero.'' On a single-node system,
communication systems are robust and relatively fast, and most systems
do not need to prepare for partial component failure. However, as
Deutsch points out, these assumptions are no longer valid in a
distributed setting. To extend Jim Gray's storage latency analogy, if
memory is around 1.5 hours from San Francisco and disk is as ``far
away'' as Pluto (2 years), in a distributed context, another computer
can be even farther---nearing or even worse than Gray's location of
tape access near Andromeda (2,000 years)---with the added concern that
messages can be lost in transit, destroyed by ``black holes'' in the
form of network partitions. Accordingly, accounting for partial
failure and communication expense is core to distributed systems and
algorithm design; ignoring them ``cause[s] big trouble and painful
learning experiences.''

@@

We summarize the results of our network measurement study in
Table~\ref{table:rtt}. On average, inter-datacenter communication is
between 1.82 and 6.38 times faster than across geographically
co-located datacenters and 40 and 647 times faster than across
geographically distributed datacenters. The cost of wide-area
communication exceeds the speed of light: for example, communicating
from S\~{a}o Paulo to Singapore is lower-bounded by speed-of-light
transit RTT of 106.7 ms but, on average, incurs a 362.8 ms RTT (95th
percentile: 649ms). This is perhaps unsurprising due to modern
Internet topologies, but this three-fold overhead is
non-negligible. As shown in Figure~\ref{fig:rtt}, the distribution of
latencies varies between links, but the trend is clear: remote
communication has a substantial cost.

@@

* As an example of a real system going from "CP" to "AP", [PNUTS dropped per-record mastering](\url{http://developer.yahoo.com/blogs/ydn/posts/2010/06/sherpa_update/#4})

What about existing systems?
* Database technology developed for single-node systems; gold standard: serializability
	* Serializability is not actually highly available; give an example!
	* Much of the database literature presumes serializability and does not consider high availability.
	* For the literature that doesn't presume serializability, it's not presented in a HA context?
* Consequence: traditional systems are not optimized for high availability
	* Consider two-phase locking in a distributed environment
	* As we will see (forward reference to Evaluation), many existing transactional systems encounter similar difficulties
* Key question in this paper: What transactional semantics are highly available, and which aren't?
	* Side note: there are infinite incomparable consistency models (e.g., always return 2, always return 3, …)
	* Our goal is to unify distributed systems literature with ACID transaction model.
* Argument: Highly Available Transactional Systems (HATS) require rethinking existing models. This work is a first step.

In this sense, HATs are similar to RAID: optimizing for graceful
handling of a worst-case failure scenario improves average-case
performance.

>>>>

Peter Deutsch begins his list of ``Fallacies of Distributed
Computing'' with two concerns: ``\textit{1.)}  The network is
reliable. \textit{2.)} Latency is zero''~\cite{fallacies-deutsch}. On
a single-node system, communication systems are robust and relatively
fast, and most systems do not need to prepare for partial component
failure. However, as Deutsch points out, these assumptions are no
longer valid in a distributed setting. To extend Jim Gray's storage
latency analogy, if memory is the distance from San Diego to Olympia
(1.5 hours), and disk is as ``far away'' as Pluto (2 years), in a
distributed context, another computer can be even farther---nearing or
even worse than Gray's location of tape access near Andromeda (2,000
years)---with the added concern that messages can be lost in transit,
destroyed by ``black holes'' in the form of network
partitions. Accordingly, accounting for partial failure and
communication expense is core to distributed systems and algorithm
design; ignoring them ``cause[s] big trouble and painful learning
experiences''~\cite{fallacies-deutsch}.

@@

We summarize the results of our network measurement study in
Table~\ref{table:rtt}. On average, inter-datacenter communication is
between 1.82 and 6.38 times faster than across geographically
co-located datacenters and 40 and 647 times faster than across
geographically distributed datacenters. The cost of wide-area
communication exceeds the speed of light: for example, communicating
from S\~{a}o Paulo to Singapore is lower-bounded by speed-of-light
transit RTT of 106.7 ms but, on average, incurs a 362.8 ms RTT (95th
percentile: 649ms). This is perhaps unsurprising due to modern
Internet topologies, but this three-fold overhead is
non-negligible. As shown in Figure~\ref{fig:rtt}, the distribution of
latencies varies between links, but the trend---in line with other
recent studies~\cite{redblue, mdcc}---is clear: remote communication
has a substantial cost.

@@
